---
title: "Cluster by Individual Samples"
author: "WYashar & TLusardi"
date: "02/02/2021"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### Set-Up

#### Libraries 

```{r setup_libs}
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)
library(clustree)
library(circlize)
library(ComplexHeatmap)
library(ggthemes)
library(viridis)
```

#### Varibles

##### Update according to your dataset

```{r setup_variables}

# Samples to process named by their treatment conditions
samples2process = c("dmso", "quzi", "ory", "combo")

# Name of the single cell project
projectName <- "flt3_aml"

# Date that Seurat object with pre-processed data was generated
importDate <- "2021-02-16"

# Clustering Resolution
clusRes <- list()
clusRes[[samples2process[1]]] <- 0.9
clusRes[[samples2process[2]]] <- 0.8
clusRes[[samples2process[3]]] <- 1.0
clusRes[[samples2process[4]]] <- 0.6
clusRes[['integrated']] <- 0.5

# Number of dimensions to reduce on
nDims <- 150
```

##### Inherent Variables

```{r setup_inherent_variables}

# Specify directory paths
directory = list(raw = "./data/raw",
                 rda = "./data/rda")

# Load Seurat objects
exptsList <- readRDS(sprintf("%s/integrated.%s.%s.rds", directory$rda, projectName,importDate))

# Clustered variables list
clusExptsList <- list()
```

### Identify UMAP Clustering Variables

``` {r id_umap_vars}
# Perform linear dimensional reduction (UMAP) on myso object
plotList <- list()

for (sample in names(exptsList)) {
  
  myso <- exptsList[[sample]]
  
  sampleAssay <- ifelse(sample=='integrated', "Integrated_snn_res." ,"SCT_snn_res.")
  
  # Create plot that explains variance by Principle Components (PCs) to find the dimensionality
  percentExplained <- function( seuratObject ){
     return(cumsum(seuratObject@reductions$pca@stdev^2)/seuratObject@reductions$pca@misc$total.variance)
  }
  
  plotPercentExplained <- function( seuratObject , ndims = nDims) {
      data.use <- percentExplained( seuratObject )
      plot <- ggplot(data = data.frame(dims = 1:ndims, var = data.use[1:ndims])) +
              geom_point(mapping = aes_string(x = 'dims', y = 'var')) + labs( y = "% variance explained", x = "PC") +
              ggtitle(sprintf("Variance Explained by Principle Components in %s", sample)) +
              theme_clean()
      return(plot)
  }
  
  plotList[[sample]][["plotPercentExplained"]] <- plotPercentExplained(myso)
  
  # Cluster the cells over a range of resolutions
  myso <- FindNeighbors(object = myso, dims = 1:nDims, verbose = FALSE)
  myso <- myso %>%
    FindClusters(resolution = seq(0,2,0.1), verbose = FALSE) %>%
    RunUMAP(dims = 1:nDims, verbose = FALSE)
  
  # Add UMAP embedding
  embeddings <- Embeddings(myso)
  myso@meta.data$PC_1 <- embeddings[,1]
  myso@meta.data$PC_2 <- embeddings[,2]
  
  # Clustree visualizations to determine best cluster resolution
  plotList[[sample]][["clustree"]] <- clustree(myso, prefix = sampleAssay) 
  
  plotList[[sample]][["clustreeOverlay"]] <- clustree_overlay(x = myso[[]],
                   prefix = sampleAssay,
                   x_value = "PC_1",
                   y_value = "PC_2") 
  
}
```

### UMAP Variable Plots {.tabset}

`r i <- 1`
#### `r toupper(names(exptsList)[i])`

##### Plot Percent Explained
```{r}
plot(plotList[[names(exptsList)[i]]][["plotPercentExplained"]])
```

##### Clustree
```{r}
plotList[[names(exptsList)[i]]][["clustree"]]
```

##### Clustree Overlay
```{r}
plotList[[names(exptsList)[i]]][["clustreeOverlay"]]
```

`r i <- 2`
#### `r toupper(names(exptsList)[i])`

##### Plot Percent Explained
```{r}
plot(plotList[[names(exptsList)[i]]][["plotPercentExplained"]])
```

##### Clustree
```{r}
plotList[[names(exptsList)[i]]][["clustree"]]
```

##### Clustree Overlay
```{r}
plotList[[names(exptsList)[i]]][["clustreeOverlay"]]
```

`r i <- 3`
#### `r toupper(names(exptsList)[i])`

##### Plot Percent Explained
```{r}
plot(plotList[[names(exptsList)[i]]][["plotPercentExplained"]])
```

##### Clustree
```{r}
plotList[[names(exptsList)[i]]][["clustree"]]
```

##### Clustree Overlay
```{r}
plotList[[names(exptsList)[i]]][["clustreeOverlay"]]
```

`r i <- 4`
#### `r toupper(names(exptsList)[i])`

##### Plot Percent Explained
```{r}
plot(plotList[[names(exptsList)[i]]][["plotPercentExplained"]])
```

##### Clustree
```{r}
plotList[[names(exptsList)[i]]][["clustree"]]
```

##### Clustree Overlay
```{r}
plotList[[names(exptsList)[i]]][["clustreeOverlay"]]
```

`r i <- 5`
#### `r toupper(names(exptsList)[i])`

##### Plot Percent Explained
```{r}
plot(plotList[[names(exptsList)[i]]][["plotPercentExplained"]])
```

##### Clustree
```{r}
plotList[[names(exptsList)[i]]][["clustree"]]
```

##### Clustree Overlay
```{r}
plotList[[names(exptsList)[i]]][["clustreeOverlay"]]
```

### UMAP Clustering by Individual Samples 

```{r umap}
plostList <- list()

for (sample in names(exptsList)) {
  
  paste(sprintf("UMAP Clustering the %s object", sample))
  
  myso <- exptsList[[sample]]
  
  # Cluster the cells over a range of resolutions
  myso <- myso %>%
    FindNeighbors(dims = 1:nDims, verbose = FALSE) %>%
    FindClusters(resolution = clusRes[[sample]], verbose = FALSE) %>%
    RunUMAP(dims = 1:nDims, verbose = FALSE)
  
  # Plot and save UMAP clusters
  Idents(myso) <- myso$seurat_clusters
  plotList[[sample]][["umap"]] <- DimPlot(object = myso, reduction = 'umap', label = TRUE, label.size = 7) 
  
  # Save clustering
  clusExptsList[[sample]] <- myso
}
```

### UMAP Plots {.tabset}

`r i <- 1`
#### `r toupper(names(exptsList)[i])`

##### UMAP 

```{r}
plotList[[names(exptsList)[i]]][["umap"]]
```

`r i <- 2`
#### `r toupper(names(exptsList)[i])`

##### UMAP 

```{r}
plotList[[names(exptsList)[i]]][["umap"]]
```

`r i <- 3`
#### `r toupper(names(exptsList)[i])`

##### UMAP 

```{r}
plotList[[names(exptsList)[i]]][["umap"]]
```

`r i <- 4`
#### `r toupper(names(exptsList)[i])`

##### UMAP 

```{r}
plotList[[names(exptsList)[i]]][["umap"]]
```

`r i <- 5`
#### `r toupper(names(exptsList)[i])`

##### UMAP 

```{r}
plotList[[names(exptsList)[i]]][["umap"]]
```

### Save Data

```{r save_data}
file2save <- sprintf("clustered.%s.%s.rds", projectName, Sys.Date())
print(sprintf("Saving clustered data by individual samples in %s", file2save))
saveRDS(clusExptsList, file = paste(directory$rda, file2save, sep = "/"))
```


### Session Information

```{r session_info}
sessionInfo()
```