---
date: "`r Sys.Date()`"
params:
  input_rds:
    value: x
  output_rds:
    value: x
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

In order to integrate all CITE-Seq datasets together, the RNA (SCT normalized) and ADT (CLR normalized) from all samples will be [integrated separately](https://github.com/satijalab/seurat/issues/2706).
These two objects will be [integrated together](https://github.com/satijalab/seurat/issues/3890#issuecomment-787598143) to produce one Seurat object that contains all samples and all their modalities.

# Set-Up

## Libraries 

```{r setup_libs, include=TRUE}
library(Seurat)
library(yaml)
library(future)
library(dplyr)
```

# Variables

## Update according to your dataset

```{r setup_variables, include=TRUE}

# read config file in yaml format
if (!file.exists("config.yaml")) {
  stop("config.yaml does not exist. Exiting the program")
} else {
  config_file <- read_yaml("config.yaml")
}

# Samples to process named by their treatment conditions
samples2process <- config_file$samples2process

# Set the baseline condition
baseline <- config_file$baseline

# Number of dimensions to reduce on
nDims <- config_file$integration_anchor_PC

# how many neighbors (k) to use when picking anchors
k.anchor <- config_file$`k.anchor`

# List of antibodies in CITE-Seq assay
ab_list <- config_file$antibody_list

doc_title <- paste(config_file$title, "- integration")
author_list <- paste(config_file$authors, collapse = ", ")
```

---
title: "`r doc_title`"
author: "`r author_list`"
---

## Inherent Variables

```{r setup_inherent_variables, include=TRUE}

# Load Seurat objects
exptsList <- readRDS(params$input_rds)

# Needed to avoid error in getGlobalsandPackges 
options(future.globals.maxSize = 16 * 1024^3)
```

# Integration - Transcriptome

```{r integrate, include=TRUE}

# Collect non-redudant list of features across all samples
rnaFeatures <- NULL

for (sample in names(exptsList)) {
  DefaultAssay(exptsList[[sample]]) <- 'SCT'
  rnaFeatures <- unique(  c(rnaFeatures, rownames(exptsList[[sample]]))  )
}

# Select most variable features for integration
intFeatures <- SelectIntegrationFeatures(object.list = exptsList,
                                        nfeatures = length(rnaFeatures),
                                        fvf.nfeatures = rnaFeatures, 
                                        assay = rep( "SCT", length(samples2process)) )

# Calculate Pearson Residuals
preppedExptsList <- PrepSCTIntegration(object.list = exptsList,
                              anchor.features = intFeatures,
                              verbose = FALSE, 
                              assay = rep( "SCT", length(samples2process)) )

# Identify integration anchors
ref <- which(names(exptsList) == baseline)
intAnchors <- FindIntegrationAnchors(object.list = preppedExptsList,
                                    normalization.method = "SCT",
                                    k.anchor = k.anchor,
                                    assay = rep("SCT", length(samples2process)),
                                    reference = ref, 
                                    dims = 1:nDims,
                                    anchor.features = intFeatures,
                                    verbose = TRUE)

# Integrate selected data
integratedRNA <- IntegrateData(anchorset = intAnchors, 
                              normalization.method = "SCT",
                              dims = 1:nDims,
                              verbose = TRUE, 
                              new.assay.name = "integrated")

# Run PCA on the integrated object
integratedRNA <- RunPCA(integratedRNA, npcs = nDims, verbose = FALSE)

```

## Scree Plot - Transcriptome {.tabset}

```{r elbowPlots, results = 'asis'}
for (i in names(exptsList)) {
  cat(paste("##", toupper(i), "\n\n"))

  cat("\n\n")
  plot(ElbowPlot(exptsList[[i]], ndims = nDims))
  cat("\n\n")
}
```

# Integration - ADT

```{r integrateADT}

adtFeatures <- NULL

for (sample in samples2process) {
  DefaultAssay(exptsList[[sample]]) <- "ADT"
  adtFeatures <- unique(  c(adtFeatures, rownames(exptsList[[sample]]))  )
  exptsList[[sample]] = exptsList[[sample]] %>% RunPCA(npcs = length(ab_list) - 1) # PCA for rPCA integration.
}

adtDims <- length(ab_list) - 1

# Use reciprocal PCA to integrate ADT

intFeatures <- SelectIntegrationFeatures(object.list = exptsList,
                                         nfeatures = length(adtFeatures),
                                         fvf.nfeatures = adtFeatures, 
                                         assay = rep( "ADT", length(samples2process)) )

print("Integrating all samples using the following ADT features...")
print(intFeatures)

ref <- which(names(exptsList) == baseline)
anchors <- FindIntegrationAnchors(object.list = exptsList,
                                  reference = ref,
                                  normalization.method = "LogNormalize",
                                  reduction = "rpca",
                                  dims = 1:adtDims,
                                  assay = rep( "ADT", length(samples2process)) )
integratedADT <- IntegrateData(anchorset = anchors, dims = 1:adtDims)

# Run PCA on the integrated ADT object
integratedADT <- integratedADT %>% ScaleData() %>% FindVariableFeatures() %>% RunPCA(npcs = adtDims, verbose = FALSE)

# Source: https://github.com/satijalab/seurat/issues/3890#issuecomment-756885252

```

# Generate integrated object

```{r integratedObject}

# add integrated ADT information to integratedRNA object.
integratedRNA[["iadt"]] <- integratedADT[["integrated"]]
integratedRNA[["pca.adt"]] <- integratedADT[["pca"]]

# WNN + UMAP
integratedRNA <- FindMultiModalNeighbors(
  integratedRNA,
  reduction.list = list("pca", "pca.adt"),
  dims.list = list(1:nDims, 1:adtDims),
  modality.weight.name = "RNA.weight"
)

# integratedRNA <- RunUMAP(integratedRNA, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")
# integratedRNA <- FindClusters(integratedRNA, graph.name = "wsnn", algorithm = 3, resolution = 2, verbose = FALSE)

exptsList[['integrated']] <- integratedRNA

```

# Save Data
```{r save_data, include=TRUE}
print(sprintf("Saving preprocessed individual samples and the integrated object in %s", params$output_rds))
saveRDS(exptsList, file = params$output_rds)
```

# Session Information
```{r session_info, include=TRUE}
sessionInfo()
```