---
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6)
```

# Set-Up

## Libraries

```{r setup_libs}
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)
library(clustree)
library(ggthemes)
library(viridis)
library(yaml)
library(tibble)
library(future)
library(plotly)
library(RColorBrewer)
library(parallel)
```

## Variables

## Update according to your dataset

```{r setup_variables}

# stderr to log file
f <- file("logs/cluster_report.err", "w")
sink(f, type = "message")

# Name of the single cell project
projectName <- snakemake@config$projectName

doc_title <- paste(snakemake@config$title, "- clustering")
author_list <- paste(snakemake@config$authors, collapse = ", ")

# Clustering Resolution

clusRes <- lapply(names(snakemake@config$resolutions), function(x) {snakemake@config$resolutions[[x]]}) # resolution values (floats)
names(clusRes) <- names(snakemake@config$resolutions) # map sample name to resolution

# Number of dimensions to reduce on
nDims <- snakemake@config$nDims

# check if resolutions > 2
if (any(clusRes > 2)) {
  stop("Cluster resolution is > 2")
}

# define cores
plan("multiprocess", workers = snakemake@threads) # no need to set seed despite random number generator warnings.

# Create plot that explains variance by Principle Components (PCs) to find the dimensionality
percentExplained <- function( seuratObject ){
  return(cumsum(seuratObject@reductions$pca@stdev^2)/seuratObject@reductions$pca@misc$total.variance)
}

plotPercentExplained <- function( seuratObject , ndims = nDims) {
    data.use <- percentExplained( seuratObject )
    plot <- ggplot(data = data.frame(dims = 1:ndims, var = data.use[1:ndims])) +
            geom_point(mapping = aes_string(x = 'dims', y = 'var')) + labs( y = "% variance explained", x = "PC") +
            ggtitle(sprintf("Variance Explained by Principle Components in %s", sample)) +
            theme_clean()
    return(plot)
}

# Find all DE genes for each cluster. This will take a long time.
doFindAllMarkers <- snakemake@config$findAllMarkers

```

---
title: "`r doc_title`"
author: "`r author_list`"
---

## Inherent Variables

```{r setup_inherent_variables}

# Load Seurat objects
exptsList <- readRDS(snakemake@input[[1]])

```

# Identify UMAP Clustering Variables

``` {r id_umap_vars}
# Perform linear dimensional reduction (UMAP) on myso object
plotList <- list()

for (sample in names(exptsList)) {
  
  myso <- exptsList[[sample]]
  
  sampleAssay <- ifelse(sample=='integrated', "Integrated_snn_res." ,"SCT_snn_res.")
  
  plotList[[sample]][["plotPercentExplained"]] <- plotPercentExplained(myso, ndims = nDims)

  plotList[[sample]][["ElbowPlot"]] <- ElbowPlot(myso, ndims = nDims) +
      ggtitle(paste("SD Explained by Principle Components in", sample)) +
      theme_clean()

  # Cluster the cells over a range of resolutions
  myso <- FindNeighbors(object = myso, dims = 1:nDims, verbose = FALSE)
  myso <- myso %>%
    FindClusters(resolution = seq(0.1, 2, 0.1), verbose = FALSE) %>%
    RunUMAP(dims = 1:nDims, verbose = FALSE)
  
  # Add UMAP embedding
  embeddings <- Embeddings(myso)
  myso@meta.data$PC_1 <- embeddings[,1]
  myso@meta.data$PC_2 <- embeddings[,2]
  
  # Clustree visualizations to determine best cluster resolution
  plotList[[sample]][["clustree"]] <- clustree(myso, prefix = sampleAssay) 
  
  plotList[[sample]][["clustreeOverlay"]] <- clustree_overlay(x = myso[[]],
                   prefix = sampleAssay,
                   x_value = "PC_1",
                   y_value = "PC_2") 
  
}

rm(myso)

plan("sequential")
```

# UMAP Parameters per Sample {.tabset}

```{r results='asis'}
for (i in names(exptsList)) {
  cat(paste( "##", toupper(i), "\n\n" ))

  cat("### Percent Variance Explained by PC\n")
  plot(plotList[[i]][["plotPercentExplained"]])
  cat("\n\n---\n\n")

  cat("### SD Explained by PC\n")
  plot(plotList[[i]][["ElbowPlot"]])
  cat("\n\n---\n\n")

  cat("### Clustree Hierarchy\n")
  plot(plotList[[i]][["clustree"]])
  cat("\n\n---\n\n")
 
  cat("### Clustree Overlay\n")
  plot(plotList[[i]][["clustreeOverlay"]])
  cat("\n\n---\n\n")
}
```

# UMAP Clustering per Sample

```{r umap}

for (sample in names(exptsList)) {
  
  paste(sprintf("UMAP Clustering the %s object", sample))
  
  myso <- exptsList[[sample]]
  
  # Cluster the cells over a range of resolutions
  myso <- myso %>%
    FindNeighbors(dims = 1:nDims, verbose = FALSE) %>%
    FindClusters(resolution = clusRes[[sample]], verbose = FALSE) %>%
    RunUMAP(dims = 1:nDims, verbose = FALSE, n.components = 3)
  
  # Plot and save UMAP clusters
  Idents(myso) <- myso$seurat_clusters
  plotList[[sample]][["umap"]] <- DimPlot(object = myso, reduction = 'umap', label = TRUE, label.size = 7) 
  
  # Save clustering
  exptsList[[sample]] <- myso
}

rm(myso)
```

# UMAP Plots {.tabset}

```{r results='asis'}
for (i in names(exptsList)) {
  cat(paste( "##", toupper(i), "\n" ))

  cat("\n")
  plot(plotList[[i]][["umap"]])
  cat("\n\n")
}
```

# ADT per Cluster - FeaturePlots {.tabset}
```{r adtNested, results='asis'}
# nested tabs. 
# tab1 = samples
# tab2 = set of antibodies
ab_list <- rownames(exptsList[[1]]@assays$ADT@counts)

for (sample in names(exptsList)) {
  cat(paste("##", toupper(sample), "{.tabset}", "\n\n"))

  for (ab in ab_list) {
    cat(paste("###", ab, "\n\n"))

    cat("\n\n")
    plot(FeaturePlot( exptsList[[sample]], ab , pt.size = 1, label = TRUE))
    cat("\n\n")

  }
}

# export ADT all markers in one plot
all.adt <- lapply(names(exptsList), function(x) {
  p <- FeaturePlot(exptsList[[x]], ab_list, pt.size = 1, label = TRUE) + 
    plot_annotation(title = paste("Antibody Markers per Cluster for", x))
  out_file <- paste0("data/markers/ADT-per-cluster-", x, ".png")
  # ggsave(out_file, p, device = "png", width = 16, height = 9, units = "in", dpi = 600)
  p
})
names(all.adt) <- names(exptsList)
```

# ADT per Cluster - Violin Plots {.tabset}
```{r results='asis'}
for (i in names(exptsList)) {
  cat(paste("##", toupper(i), "\n\n"))
  # Wrapped violin plots of ADT per cluster per sample. Fetch metadata first.
  metadata.adt <- merge(t(as.matrix(exptsList[[i]]@assays$ADT@counts)), exptsList[[i]]@meta.data, by = 0)
  metadata.adt <- metadata.adt %>% 
                      rename("cell" = "Row.names") %>% 
                      select(cell, matches("*ab-*"), seurat_clusters)
  metadata.adt.melt <- reshape2::melt(metadata.adt)
  p <- ggplot(metadata.adt.melt, aes(x = variable, y = log(value + 1), fill = variable, color = variable)) +
    geom_violin() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    facet_wrap(. ~ seurat_clusters)
  cat("\n\n")
  print(p)
  cat("\n\n")
}
```

# 3D UMAP Colored by Clusters

```{r}

# define umap output directory
if (!dir.exists("data/umaps")) {
  dir.create("data/umaps")
}

for (i in names(exptsList)) {

  myso <- exptsList[[i]]

  cell_meta <- merge(myso@reductions$umap@cell.embeddings, myso@meta.data, by = 0, all = TRUE)
  cell_meta <- cell_meta %>% column_to_rownames("Row.names")
  col_scheme <- colorRampPalette(brewer.pal(8, snakemake@config$rcolorbrewer_palette))(length(unique(cell_meta$seurat_clusters)))

  p <- plot_ly(cell_meta,
          x = ~UMAP_1,
          y = ~UMAP_2,
          z = ~UMAP_3,
          size = 1,
          color = ~seurat_clusters,
          colors = col_scheme,
          # hover text
          text = ~paste("Cluster:", seurat_clusters, "<br>nFeature_ADT:", nFeature_ADT, "<br>nCount_SCT:", nCount_SCT)) %>% 
    add_markers() %>%
    layout(title = paste(toupper(i), "UMAP Clusters"),
           xaxis = list(title = "UMAP_1"),
           yaxis = list(title = "UMAP_2"),
           zaxis = list(title = "UMAP_3"))

  # save and print where they are located.
  out_file <- paste0("data/umaps/", tolower(i), "_umap.html")
  print(paste("Sample", i, "UMAP file is located at:", out_file))
  # htmlwidgets::saveWidget(p, out_file) # Rmd does not output files.
}

rm(myso)
```

# 3D UMAP UMAP Colored by Identity

```{r}

if (!dir.exists("data/umaps")) {
  dir.create("data/umaps")
}

myso <- exptsList[["integrated"]]

cell_meta <- merge(myso@reductions$umap@cell.embeddings, myso@meta.data, by = 0, all = TRUE)
cell_meta <- cell_meta %>% column_to_rownames("Row.names")
col_scheme <- colorRampPalette(brewer.pal(8, snakemake@config$rcolorbrewer_palette))(length(unique(cell_meta$seurat_clusters)))

p <- plot_ly(cell_meta,
        x = ~UMAP_1,
        y = ~UMAP_2,
        z = ~UMAP_3,
        size =1,
        color = ~orig.ident,
        colors = col_scheme,
        # hover text
        text = ~paste("Cluster:", seurat_clusters, "<br>nFeature_ADT:", nFeature_ADT, "<br>nCount_SCT:", nCount_SCT)) %>% 
  add_markers() %>%
  layout(title = "Integrated UMAP by Sample Identity",
         xaxis = list(title = "UMAP_1"),
         yaxis = list(title = "UMAP_2"),
         zaxis = list(title = "UMAP_3"))

out_file <- "data/umaps/integrated_identity_umap.html"
print(paste("Integrated UMAP colored by sample identity is here:", out_file))
htmlwidgets::saveWidget(p, out_file)

rm(cell_meta, col_scheme, p, myso, out_file)

```

```{r cluster_markers}

# parallelize FindAllMarkers. Share snakemake@config$cores across all samples.
if (doFindAllMarkers) {

  if (!dir.exists("data/markers")) {
  dir.create("data/markers")
  }

  allSampleMarkers <- mclapply(names(exptsList), function(x) {
    a <- FindAllMarkers(exptsList[[x]])
    a
  }, mc.cores = floor(snakemake@config$cores / length(names(exptsList))) )

  names(allSampleMarkers) <- names(exptsList)
}
# mclapply returns results in the right order. https://stackoverflow.com/questions/14697901/is-mclapply-guaranteed-to-return-its-results-in-order

# make plots and export tables
if (doFindAllMarkers) {
  plotList <- list()

  for (sample in names(allSampleMarkers)) {
    # j is all DE genes per cluster for an individual sample
    j <- allSampleMarkers[[sample]]
    top <- j %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC)
    plotList[[sample]][['topMarkerHeatmap']] <- DoHeatmap(exptsList[[sample]], features = top$gene) + NoLegend()
    # export allMarkers for each sample and heatmap
    # write.table(j, file = paste0("data/markers/", sample, "-markers.txt"), row.names = TRUE, sep = "\t", quote = FALSE)
    # png(paste0("data/markers/", sample, "_heatmap.png"), width = 16, height = 9, units = "in", res = 300)
    # plotList[[sample]][['topMarkerHeatmap']]
    # dev.off()
  }
}
```

```{r results='asis'}
# format the output
if (doFindAllMarkers) {
  cat("\n\n# Top Cluster Markers {.tabset}\n\n")
  for (sample in names(exptsList)) {
    cat(paste("\n\n##", toupper(sample), "\n\n"))
    cat("\n\n")
    print(  plotList[[sample]][['topMarkerHeatmap']] + theme(text = element_text(size = 4, face="bold"))  )
    cat("\n\n")
  }
}
```

# Session Information

```{r session_info}
sessionInfo()
```